An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-17 DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
a bit new in experimental section, added one idea for conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The multiobjective risk-equity constrained routing problem}
\label{S_RECRP}

@@

\subsection{The optimization model} \label{SS_AOM}
We introduce the integer variable $y_{ij}^c$ for the number of trucks that use arc $(i,j)\in A$ for transporting commodity $c$. As mentioned above, we assume a fixed number of trucks for each commodity and that all trucks associated to a certain commodity have the same load. The proposed model is defined as follows:
\begin{eqnarray}
   \min  & \sum_{c \in C}\limits \sum_{(i,j) \in A}\limits c_{ij} y_{ij}^c & \label{eq:obj1}\\
   \min  & \sum_{c \in C}\limits \sum_{(i,j) \in A}\limits r_{ij}^{c} y_{ij}^c & \label{eq:obj2}\\
   \min  & \max_{(i,j)\in A}\limits \big\{ \sum_{c \in C}\limits r_{ij}^{c} y_{ij}^c \big\} & \label{eq:obj3}\\
    s.t. & \sum_{j \in \delta^+(i)} y_{ij}^c -  \sum_{j \in \delta^-(i)} y_{ji}^c = q_i^c\quad         & \forall i \in N,  c \in  C \label{eq:flow} \\
         & y_{ij}^c \in \{0, 1, 2, \ldots\}                                                            & \forall (i,j)\in A, c\in C\label{eq:integerY}.
\end{eqnarray}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evolutionary multiobjective optimization} \label{S_MEO}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Evolutionary algorithms (EAs) and Evolutionary Multiobjective Optimization (EMO) algorithms in particular are general-purpose randomized search heuristics and as such well suited for problems where the objective function(s) can be highly non-linear, noisy, or even given only implicitly, e.g., by expensive simulations \citep{deb2001a,cvl2007a}. Since the third objective in the above problem formulation is nonlinear, we propose to use an EMO algorithm for the multiobjective risk-equity constrained routing problem here. Our EMO algorithm follows the standard iterative/generational cycle of an EA of mating selection, variation, objective function evaluation, and environmental selection and is build upon the state-of-the-art selection scheme in HypE \citep{bz2011a} as implemented in the PISA platform \citep{bltz2003a}. The variation operators as well as the representation of the solutions, however, have to be adapted to the problem at hand in the following way in order to fulfill the problem's constraints at all times\footnote{The source code of the corresponding PISA hazmat variator is going to be available on the official PISA web page \url{http://www.tik.ee.ethz.ch/pisa/} and until then can be downloaded on the preliminary web page \url{http://researchers.lille.inria.fr/~brockhof/downloads/hazmatvariator20111110.zip}}. Moreover, two different initialization schemes are proposed to deal with the specific variable length representation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following experiments have two goals:
- show that the algorithm works in principle on practical examples in manageable time
- check which of the two initialization schemes is 
In addition, a few more experiments show that the standard parameters of the algorithm behaves as with other test problems: population size, environmental selection (change from HypE to NSGA-II)

As a test case, we chose to use a real-world case study based on the work of \citet{bianco09}. The network used here consists of 311 nodes and 441 arcs and represents the main transport roads in Italy's Lazio region which amongst other cities includes the capital Rome.

The risks $r_{ij}^c$ for each arc $(i,j) \in A$ and commodity $c \in C$ are evaluated as the societal risk computed as the number of people living inside the exposure zone around link $(i,j)$ (whose size depends on the hazmat type $c$ times the accident probability involving the vehicle), for details see  \citep{bianco09}.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Multiobjective Risk-Equity Constrained Routing Problem} \label{S_RECRP}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{The optimization model} \label{SS_AOM}
We introduce the integer variable $y_{ij}^c$ for the number of trucks that use arc $(i,j)\in A$ for transporting commodity $c$. As mentioned above, we assume a fixed number of trucks for each commodity and that all trucks associated to a certain commodity have the same load. The proposed model is defined as follows:
\begin{eqnarray}
   \min  & \sum_{c \in C}\limits \sum_{(i,j) \in A}\limits c_{ij} y_{ij}^c & \label{eq:obj1}\\
   \min  & \sum_{c \in C}\limits \sum_{(i,j) \in A}\limits r_{ij}^{c} y_{ij}^c & \label{eq:obj2}\\
   \min  & \max_{(i,j)\in A}\limits \big\{ \sum_{c \in C}\limits r_{ij}^{c} y_{ij}^c \big\} & \label{eq:obj3}\\
    s.t. & \sum_{j \in \delta^+(i)}\limits y_{ij}^c -  \sum_{j \in \delta^-(i)}\limits y_{ji}^c = q_i^c\quad         & \forall i \in N,  c \in  C \label{eq:flow} \\
         & y_{ij}^c \in \{0, 1, 2, \ldots\}                                                            & \forall (i,j)\in A, c\in C\label{eq:integerY}.
\end{eqnarray}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An Evolutionary Multiobjective Optimization Algorithm for Hazmat Routing} \label{S_MEO}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Evolutionary algorithms (EAs) and Evolutionary Multiobjective Optimization (EMO) algorithms in particular are general-purpose randomized search heuristics and as such well suited for problems where the objective function(s) can be highly non-linear, noisy, or even given only implicitly, e.g., by expensive simulations \citep{deb2001a,cvl2007a}. Since the third objective in the above problem formulation is nonlinear, we propose to use an EMO algorithm for the multiobjective risk-equity constrained routing problem here. Our EMO algorithm follows the standard iterative/generational cycle of an EA of mating selection, variation, objective function evaluation, and environmental selection and is build upon the state-of-the-art selection scheme in HypE \citep{bz2011a} as implemented in the PISA platform \citep{bltz2003a}. The variation operators as well as the representation of the solutions, however, have to be adapted to the problem at hand in the following way in order to fulfill the problem's constraints at all times\footnote{The source code of the corresponding PISA hazmat variator is going to be available on the official PISA web page \url{http://www.tik.ee.ethz.ch/pisa/} and until then can be downloaded on the preliminary web page \url{http://researchers.lille.inria.fr/~brockhof/downloads/hazmatvariator20111110.zip}}. Moreover, two different initialization schemes are proposed to deal with the specific variable length representation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the remainder of this paper, we investigate how the above proposed evolutionary multiobjective optimization algorithm can cope with realistic instances of the multiobjective risk-equity constrained routing problem. The experimental analyses have two main goals. On the one hand, we show that the algorithm is efficient and can find multiple non-dominated solutions on practical examples in reasonable time. On the other hand, we investigate which of the two proposed initialization schemes is favourable and when this is the case. In addition, a few more experiments are conducted to show that the standard parameters of the algorithm behave as with other test problems. To this end, we vary the population size of the algorithm as well as its environmental selection scheme.

As a test case for all the experiments, we chose to use a real-world case study based on the work of \citet{bianco09}. The network used here consists of 311 nodes and 441 arcs and represents the main transport roads in Italy's Lazio region which, amongst other cities, includes the capital Rome. We distinguish between instances with a different number of commodities (between 2 and 4) and with different numbers of trucks, see Table~\ref{tab:instances}.

\TODO{rewrite:} The risks $r_{ij}^c$ for each arc $(i,j) \in A$ and commodity $c \in C$ are evaluated as the societal risk computed as the number of people living inside the exposure zone around link $(i,j)$ (whose size depends on the hazmat type $c$ times the accident probability involving the vehicle), for details see  \citep{bianco09}.

\begin{table}
\caption{\label{tab:instances} Parameters of the instances used in the experimental analyses.}
\begin{tabular}{lllll}
	\toprule
	name & \begin{minipage}{1cm}\#com\-mod\-i\-ties\end{minipage} & \#trucks & cost range & risk range \\ \midrule
	lazio\_2\_1 & 2 & \small 20+19         & & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\end{minipage}\\
	lazio\_2\_2 & 2 & \small 23+29         & & \\
	lazio\_3\_1 & 3 & \small 25+16+28      & & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\end{minipage}\\
	lazio\_4\_1 & 4 & \small 21+3+30+5     & \small [0.0109239, 3.93701] & \begin{minipage}[t]{3.7cm} \small [762, 274628]\\ x [610.18, 219702.6]\\ x [1120.352, 439405.2]\\ x [762, 274628]\end{minipage}\\
	\bottomrule
\end{tabular}
\end{table}

@@

In future work, one might want to look further into adding (simple) heuristics within the evoluionary algorithm in order to improve the algorithm. For example, one can observe that the solutions produced by the algorithm sometimes uses one and the same arc within a truck path, hence introducing cylces. With the three objectives cost, total and equity risk, such cycles can be easily prevented within the variation operator which might eventually result in a faster convergence of the algorithm to the Pareto front.
